---
title: "In silico R4IDs simulation"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

## Aim: establish likely performance.

# Methods

A python script was used to simulate nanopore R9 sequencing runs from published **Cammelinae** genomes (~/Documents/all_work/programming/oddjects-sandbox/R4IDs/In-silico-genome-skimming-by-args.py). This used a mutation rate of 1/20 e.g. 5%. Uniformly-selected sites for mutation were assigned a substitution (equal base frequencies), point deletion, or point insertion (homopolymer created by inserting **n**th base as **n+1**th)

These read sets (intensities 10, 100, 1000 and 10000 reads) were used to simulate both R4IDs set-up and ID resequencing runs as for the empirical sci-fest data, and analysed in the same way.

## Input reference genomes

Species | File | Mbp | # contigs
------- | ---- | --- | ---------
A. thaliana | () | 120 | 7
Capsella rubella | file:////Users/joeparker/Downloads/ANNY01.1.fsa_nt.gz | 129 | 7,067
A. halleri subsp. gemmifera | file:////Users/joeparker/Downloads/FJVB01.1.fsa_nt.gz | 196 | 2,239
A. lyrata | () | 208 | 3645
Capsella bursa-pastoris | file:////Users/joeparker/Downloads/MPGU01.1.fsa_nt.gz | 268 | 8,186 
Camelina sativa | file:////Users/joeparker/Downloads/JFZQ01.1.fsa_nt.gz file:////Users/joeparker/Downloads/JFZQ01.2.fsa_nt.gz | 641 | 37,780

## Nanopore sequencing simulation

## Analysis

```{r}
library(lattice)
in_silico = read.table('~/Documents/all_work/programming/oddjects-sandbox/R4IDs/in-silico.out.parsed.tdf',header=T,sep="\t")

```

How do stats' performance vary by species? List species in order. 

First: *rate TP: total*

```{r}
levelplot((one_way_TP + two_way_TP) / total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species")
levelplot((one_way_TP + two_way_TP) / total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-thaliana_02',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. thaliana",sub='120Mbp; 7 contigs')
levelplot((one_way_TP + two_way_TP) / total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-rubella_03',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. rubella",sub='129Mbp; 7,067 contigs')
levelplot((one_way_TP + two_way_TP) / total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-halleri_01',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. halleri",sub='196Mbp; 2,239 contigs')
levelplot((one_way_TP + two_way_TP) / total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-lyrata_04',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. lyrata",sub='208Mbp; 3,645 contigs')
levelplot((one_way_TP + two_way_TP) / total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-bursa_05',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. bursa",sub='268Mbp; 8,186 contigs')
levelplot((one_way_TP + two_way_TP) / total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-sativa_06',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. sativa",sub='641Mbp; 37,780 contigs')
```
Secondly *rate TP:FP*:
```{r}
levelplot((one_way_TP + two_way_TP) / (one_way_FP + two_way_FP) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species")
levelplot((one_way_TP + two_way_TP) / (one_way_FP + two_way_FP) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-thaliana_02',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. thaliana",sub='120Mbp; 7 contigs')
levelplot((one_way_TP + two_way_TP) / (one_way_FP + two_way_FP) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-rubella_03',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. rubella",sub='129Mbp; 7,067 contigs')
levelplot((one_way_TP + two_way_TP) / (one_way_FP + two_way_FP) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-halleri_01',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. halleri",sub='196Mbp; 2,239 contigs')
levelplot((one_way_TP + two_way_TP) / (one_way_FP + two_way_FP) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-lyrata_04',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. lyrata",sub='208Mbp; 3,645 contigs')
levelplot((one_way_TP + two_way_TP) / (one_way_FP + two_way_FP) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-bursa_05',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. bursa",sub='268Mbp; 8,186 contigs')
levelplot((one_way_TP + two_way_TP) / (one_way_FP + two_way_FP) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-sativa_06',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. sativa",sub='641Mbp; 37,780 contigs')
```

Thirdly *mean_bias*
```{r}
levelplot(log10(mean_bias) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species")
levelplot(mean_bias ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-thaliana_02',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. thaliana",sub='120Mbp; 7 contigs')
levelplot(mean_bias ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-rubella_03',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. rubella",sub='129Mbp; 7,067 contigs')
levelplot(mean_bias ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-halleri_01',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. halleri",sub='196Mbp; 2,239 contigs')
levelplot(mean_bias ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-lyrata_04',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. lyrata",sub='208Mbp; 3,645 contigs')
levelplot(mean_bias ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-bursa_05',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. bursa",sub='268Mbp; 8,186 contigs')
levelplot(mean_bias ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-sativa_06',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. sativa",sub='641Mbp; 37,780 contigs')
```

Fourth *hit rate in 2-way assignments*
```{r}
levelplot(two_way_rate ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species")
levelplot(two_way_rate ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-thaliana_02',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. thaliana",sub='120Mbp; 7 contigs')
levelplot(two_way_rate ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-rubella_03',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. rubella",sub='129Mbp; 7,067 contigs')
levelplot(two_way_rate ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-halleri_01',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. halleri",sub='196Mbp; 2,239 contigs')
levelplot(two_way_rate ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-lyrata_04',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. lyrata",sub='208Mbp; 3,645 contigs')
levelplot(two_way_rate ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-bursa_05',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. bursa",sub='268Mbp; 8,186 contigs')
levelplot(two_way_rate ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-sativa_06',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. sativa",sub='641Mbp; 37,780 contigs')
```

Fifth *hit rate in 2-way assignments, with cutoff>50*
```{r}
levelplot(two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species")
levelplot(two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-thaliana_02',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. thaliana",sub='120Mbp; 7 contigs')
levelplot(two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-rubella_03',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. rubella",sub='129Mbp; 7,067 contigs')
levelplot(two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-halleri_01',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. halleri",sub='196Mbp; 2,239 contigs')
levelplot(two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-lyrata_04',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. lyrata",sub='208Mbp; 3,645 contigs')
levelplot(two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-bursa_05',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. bursa",sub='268Mbp; 8,186 contigs')
levelplot(two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-sativa_06',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. sativa",sub='641Mbp; 37,780 contigs')
```

Finally *total hits*
```{r}
levelplot(total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species")
levelplot(total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-thaliana_02',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. thaliana",sub='120Mbp; 7 contigs')
levelplot(total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-rubella_03',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. rubella",sub='129Mbp; 7,067 contigs')
levelplot(total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-halleri_01',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. halleri",sub='196Mbp; 2,239 contigs')
levelplot(total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-lyrata_04',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. lyrata",sub='208Mbp; 3,645 contigs')
levelplot(total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-bursa_05',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. bursa",sub='268Mbp; 8,186 contigs')
levelplot(total_hits ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-sativa_06',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. sativa",sub='641Mbp; 37,780 contigs')
```

Okay, this is a little hard to interpret; now plot hit percentage (expectation) * total hits, e.g. number of expected positives:
```{r}
levelplot(total_hits * two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species")
levelplot(total_hits * two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-thaliana_02',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. thaliana",sub='120Mbp; 7 contigs')
levelplot(total_hits * two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-rubella_03',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. rubella",sub='129Mbp; 7,067 contigs')
levelplot(total_hits * two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-halleri_01',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. halleri",sub='196Mbp; 2,239 contigs')
levelplot(total_hits * two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='A-lyrata_04',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="A. lyrata",sub='208Mbp; 3,645 contigs')
levelplot(total_hits * two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-bursa_05',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. bursa",sub='268Mbp; 8,186 contigs')
levelplot(total_hits * two_way_rate_with_cutoff ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico[in_silico$TP_species=='C-sativa_06',], col.regions = heat.colors(100)[length(heat.colors(100)):1], main="C. sativa",sub='641Mbp; 37,780 contigs')
```

Formally analyse the relationship with a GLM:
```{r}
# 2-way with cutoff, by species too
model_total_positives=glm(total_hits * two_way_rate_with_cutoff ~ (sample_DB_intensity)*(species_DB_intensity)*TP_species,data=in_silico)
summary.glm(model_total_positives)
summary(aov(model_total_positives))
# two-way rate with cutoff, no species
model_total_positives_nosp=glm(total_hits * two_way_rate_with_cutoff ~ (sample_DB_intensity)*(species_DB_intensity),data=in_silico)
summary(aov(model_total_positives_nosp))
summary(model_total_positives_nosp)
# 2-way, by species too
model_total_positives=glm(total_hits * two_way_rate ~ (sample_DB_intensity)*(species_DB_intensity)*TP_species,data=in_silico)
summary(aov(model_total_positives))
# two-way rate with cutoff, no species
model_total_positives_nosp=glm(total_hits * two_way_rate ~ (sample_DB_intensity)*(species_DB_intensity),data=in_silico)
summary(aov(model_total_positives_nosp))
summary(model_total_positives_nosp)
```
And what we might really want to know is the expectation that a read in a sequencing run of length n will be a positive:
```{r}
levelplot((total_hits * two_way_rate)/sample_DB_intensity ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species: TP hit expectation")
levelplot((total_hits * two_way_rate_with_cutoff)/sample_DB_intensity ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species: TP hit expectation with cutoff")
levelplot(((total_hits * ((one_way_TP + two_way_TP)/(one_way_FP + two_way_FP+one_way_TP+two_way_TP)))/sample_DB_intensity) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species: Accuracy expectation (TP:TP+FP)/total reads")
levelplot((( ((one_way_TP + two_way_TP)/(one_way_FP + two_way_FP)))/sample_DB_intensity) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species: Accuracy")
levelplot((total_hits /sample_DB_intensity) ~ log10(sample_DB_intensity)*log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species: total hit rate (all reads)")
```

Let's break down the expected number of good hits a bit more simply:
```{r}
par(mfrow=c(1,2))
boxplot((( ((one_way_TP + two_way_TP)))/sample_DB_intensity) ~ log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 reference DB size',ylab='true hits / total reads')
boxplot((( ((one_way_TP + two_way_TP)))/sample_DB_intensity) ~ log10(sample_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 total reads size',ylab='true hits / total reads')
```

As above but using two-way hit rate:
```{r}
par(mfrow=c(1,2))
boxplot((total_hits * two_way_rate)/sample_DB_intensity ~ log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 reference DB size',ylab='two-way expected hits / total reads')
boxplot((total_hits * two_way_rate)/sample_DB_intensity ~ log10(sample_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 total reads size',ylab='two-way expected hits / total reads')
```

As above but using two-way hit rate, with cutoff > 50:
```{r}
par(mfrow=c(1,2))
boxplot((total_hits * two_way_rate_with_cutoff)/sample_DB_intensity ~ log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 reference DB size',ylab='two-way expected hits (cutoff) / total reads')
boxplot((total_hits * two_way_rate_with_cutoff)/sample_DB_intensity ~ log10(sample_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 total reads size',ylab='two-way expected hits (with cutoff) / total reads')
```

As above but accuracy e.g. [(TP:TP+FP) / total reads]. This is probably closest to the headline 'effective ID rate', e.g. expectation that a given sequenced read produces a BLAST hit which is accurately a true positive.
```{r}
par(mfrow=c(1,2))
boxplot(((total_hits * ((one_way_TP + two_way_TP)/(one_way_FP + two_way_FP+one_way_TP+two_way_TP)))/sample_DB_intensity) ~ log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 reference DB size',ylab='Accuracy expectation (TP:TP+FP)/total reads')
boxplot(((total_hits * ((one_way_TP + two_way_TP)/(one_way_FP + two_way_FP+one_way_TP+two_way_TP)))/sample_DB_intensity) ~ log10(sample_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 total reads size',ylab='Accuracy expectation (TP:TP+FP)/total reads')
```

Of course we should ask the reciprocal question e.g. [(FP:TP+FP) / total reads], expectation that a given sequenced read produces a BLAST hit which is inaccurately a false positive.
```{r}
par(mfrow=c(1,2))
boxplot(((total_hits * ((one_way_FP + two_way_FP)/(one_way_FP + two_way_FP+one_way_TP+two_way_TP)))/sample_DB_intensity) ~ log10(species_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 reference DB size',ylab='Miss expectation (FP:TP+FP)/total reads')
boxplot(((total_hits * ((one_way_FP + two_way_FP)/(one_way_FP + two_way_FP+one_way_TP+two_way_TP)))/sample_DB_intensity) ~ log10(sample_DB_intensity),data=in_silico, col.regions = heat.colors(100)[length(heat.colors(100)):1], main="All species",xlab='log10 total reads size',ylab='Miss expectation (FP:TP+FP)/total reads')
```