---
title: "Plot UK crime data"
author: '[Joe Parker/@lonelyjoeparker](https://github.com/lonelyjoeparker)'
date: '2018-01-12'
output:
  html_document: default
  html_notebook: default
---

Quick exercise in plotting UK national crime database data using OpenStreetMap, and fitting a contour (kernel density estimator) function for 2-D spatial smoothing.

## 1. Prerequisites

```{r}
# First the required packages.
# The top-level ones are OpenStreetMap, MASS and ks
# Other dependencies below this include:
#   X11, rgl, rgdal

# On linux this will mean something like:
#   sudo apt-get update && sudo apt-get install libcgal-dev libglu1-mesa-dev libglu1-mesa-dev libx11-dev gdal-bin gdal-config

# I googled the shit out of this, BTW:
#   https://stackoverflow.com/questions/15248815/rgdal-package-installation

# install.packages("OpenStreetMap")
# install.packages("ks")
# install.packages("MASS")

# load required libs
library(OpenStreetMap)
library(MASS)
library(ks)

# Get the crime data itself
# From e.g. https://wwww.police.uk/surrey/RUCM/crime/stats
# Saved to ~/Downloads/c1b1d294-4b4b-4a47-8f9c-4dcd7af9e56a.csv
crime <- read.csv("~/Downloads/c1b1d294-4b4b-4a47-8f9c-4dcd7af9e56a.csv")
head(crime)
```
The timestamps here aren't important but might as well convert them:
```{r}
crime$datmon <- as.Date(paste(crime$Month,'-01',sep=''))
```

## Plotting raster maps

Now we're ready to plot the map itself. Notes from https://www.r-bloggers.com/plot-maps-like-a-boss/. 
```{r}
# get the map
chertsey <- (openmap(c(51.40,-0.52), c(51.37,-0.48)))
plot(chertsey)
```
Looks cool! 

Note, `map()` takes a top-left then bottom-right pair, e.g. `map(c(x1,y1),c(x2,y2))` where x will be longitude and y latitude co-ords. This produces a plotting device you can add to like any other `plot()` call, *BUT* without transformation/projection the points returned in e.g. OpenStreetMap will be in (I think) metres N and E from the equator/meridien (I think). This means that to plot points on them directly you have to do some guessing:
```{r}
plot(chertsey)
abline(v=seq(-57886,0,length.out=30),col="green",lwd=4)
# wtf?
```

## 3. Co-ordinate projection and surface prediction

The lines appear at seemingly crazy intervals ("aren't these lat/long co-ordinates? What does a longitude of '-57,000' mean?!"). This is because, importantly, the crime dataset lat/long points refer to a flat 2D surface, whereas we need to project to the OpenStreetMap spheroid. 

Finally we can plot and visualise our surface:

```{r}
#transform map from OpenStreetMap co-ords to 2D lat-long projection
map_ll <- openproj(chertsey,projection="+proj=longlat")

# plot map
plot(map_ll)

# plot individual crimes on the map in big fat purple
points(Latitude~Longitude,data=crime,col="purple",pch=19,cex=1.5)

# create a contour (surface) plot using KDE
crm <- kde2d(crime$Longitude,crime$Latitude)

# add the contour
contour(crm,add=TRUE,col=rev(heat.colors(11)),lwd=5)
```
Hey presto! 

BTW we might also wonder if particular categories of crime occur in particular places. I can't be arsed to do this in full but clearly we'd subset the data, etc. Quick check with `boxplot()` tho:

```{r}
boxplot(Latitude~Category,data=crime)
boxplot(Longitude~Category,data=crime)
```

So... there aren't really any particular locations more/less prone to any type of crime.